### 2025-09-24 paper/sec/authoring-workflow.tex

\item \textbf{Initial configuration.} The author imports the target text and accompanying data into the system
to create a programmatic representation of the target document. Initially this is simply isomorphic to the
target text, taking the form of a string literal \kw{"""..."""}, where the triple quotes are Fluid syntax for
a Python or JavaScript-style \emph{interpolated string}, i.e.~a literal where expressions of the form
\kw{\{$e$\}} are permitted within the string. The \SuggestionAgent analyses the target text and identifies any
fragments which are candidates for being computed instead of remaining as literal substrings.

\item \textbf{High-level Authoring workflow.} The system then enters the human-in-the loop authoring workflow
shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
system waits for the author to select a fragment of text $s$ to interpret (perhaps previously highlighted by
the \SuggestionAgent). The system then attempts to generates a candidate Fluid expression $e$ using the
closed-loop synthesis step (3) below. If code synthesis succeeds with an expression $e$, the system proceeds
to the manual validation step (4) below. If the synthesis step fails with no expression, no remedial action is
possible; this is considered an unsuccessful path through the workflow and returns the system to the entry
state. Otherwise the synthesis step produces an expression $e$ which evaluates to a mismatched string $s' \neq
s$ outcome, and the user can choose to manually abort and return to the entry state, or optionally to
\emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the candidate
expression. This is intended to cover the situation where the author has made a claim which is
\emph{incorrect}, and the data set and surrounding natural language have led the LLM to synthesise an
expression which generates a different value from the one specified by the user.

>>>>

\item \textbf{Initial configuration.} The author imports the target text and accompanying data into the system
to create a programmatic representation of the target document. Initially this is simply isomorphic to the
target text, taking the form of a string literal \kw{"""..."""}, where the triple quotes are Fluid syntax for
a Python or JavaScript-style \emph{interpolated string}, i.e.~a literal where expressions of the form
\kw{\{$e$\}} are permitted within the string. The \SuggestionAgent analyses the target text and identifies any
fragments which are candidates for being computed instead of remaining as literal substrings.
\mdnote{Since this is the only occurrence of the term ``isomorphic'' in the paper, it might be worth adding a brief explanation, unless you feel it is clear enough as is.}

\item \textbf{High-level Authoring workflow.} The system then enters the human-in-the loop authoring workflow
shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
system waits for the author to select a fragment of text $s$ to interpret (perhaps previously highlighted by
the \SuggestionAgent). The system then attempts to generate \mdnote{I changed ``generates'' to ``generate''} a candidate Fluid expression $e$ using the
closed-loop synthesis step (3) below. If code synthesis succeeds with an expression $e$, the system proceeds
to the manual validation step (4) below. If the synthesis step fails with no expression, no remedial action is
possible; this is considered an unsuccessful path through the workflow and returns the system to the entry
state. Otherwise the synthesis step produces an expression $e$ which evaluates to a mismatched string $s' \neq
s$ outcome, and the user can choose to manually abort and return to the entry state, or optionally to
\emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the candidate
expression. This is intended to cover the situation where the author has made a claim which is
\emph{incorrect}, and the data set and surrounding natural language have led the LLM to synthesise an
expression which generates a different value from the one specified by the user.

### 2025-09-24 paper/sec/conclusion.tex
Tweaks to eval.

\section{Conclusion}
\label{sec:conclusion}

\input{fig/nl-further-works}

>>>>

\section{Conclusions and Future Work}
\label{sec:conclusion}

\subsection{Future Work}
\label{sec:conclusion:future-work}

Providing helper functions such as \kw{growShrink} and \kw{trendWord} as a predefined library is in some ways
beneficial because they offer a uniform framework for interpreting a given scholarly document. We would also
like make the system able to operate in the absence of these, perhaps adding a loopback condition to turn
``Definition not found'' errors into an augmented prompt to generate the appropriate definition.

Another important direction is to broaden the scope of the artifacts than can be referred to and the natural
language forms that can be used to refer to them. For example, many claims in scholarly writing refer to
visualisations and intermediate data sets derived via cleansing, imputation or aggregation. These should also
be made available to the LLM (for visualisations, as structured data rather than as static bitmaps), enabling
interpretation of natural language in terms of visual information and other intermediate analyses. Extending
coverage to additional idioms beyond those considered in Table~\ref{tab:natural-language-forms}, such as
cardinals and multiplicatives (\textfrag{three} and \textfrag{three times}), rounding (\textfrag{about 2
degrees warmer}) and graded adjectives \textfrag{extremely likely}, would also expand the expressivity of the
assistant.

Consider the false statement \textfrag{The population of France is the largest in the EU}. There are at least
two ways to repair this statement so that it comes out true: treat the subject (France) as fixed, but adjust
the predicate so that it matches the facts (\textfrag{second largest}), or treat the predicate as fixed, but
keep adjust the subject to match (\textfrag{Germany}). To make these sorts of (pragmatically important)
distinctions, we may need to introduce the idea of \emph{referential terms} that have fixed denotations, in
contrast to queries, whose values are contingent on data.

Finally, integration into developer and authoring environments such as VSCode or Cursor would make the
workflow more seamless for authors. Automatic generation of the counterfactual test cases discussed in
\secref{validation-with-counterfactual-data} could further strengthen validation at authoring-time, improving
the trustability of interpreted claims.

### 2025-09-24 paper/sec/introduction.tex

\begin{figure}%[h]
    \centering
    \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-with-pointer.png}
    \vspace{1mm}
    \hrule
    \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-counterfactual-with-pointer.png}
    \caption{\emph{Transparent document} example, showing data-driven natural language (top
    vs.~bottom)}
    \label{fig:scigen-example-website}
\end{figure}

>>>>

\begin{figure}%[h]
    \centering
    \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-with-pointer.png}
    \vspace{1mm}
    \hrule
    \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-counterfactual-with-pointer.png}
    \caption{\emph{Transparent document} example, showing data-driven natural language (top
    vs.~bottom) \mdnote{The caption could perhaps say a little more about the figure, although it is already described in the text.}}
    \label{fig:scigen-example-website}
\end{figure}
